A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2023-07-12 kobakian-cook.tex
Review comments addressed

\abstract{%
Spatial distributions have been presented on alternative representations
of geography, such as cartograms, for many years. In modern times,
interactivity and animation have allowed alternative displays to play a
larger role. Alternative representations have been popularised by online
news sites, and digital atlases with a focus on public consumption.
Applications are increasingly widespread, especially in the areas of
disease mapping, and election results. The algorithm presented creates a
display that uses tessellated hexagons to represent a set of spatial
polygons, and is implemented in the R package called sugarbag. It
allocates these hexagon in a manner that preserves the spatial
relationship of the geographic units, in light of their positions to
points of interest. The display showcases spatial distributions, by
emphasising the small geographical regions that are often difficult to
locate on geographic maps.
}

@@

Over time, the density of living in major cities has increased as
residents have gathered to live near urban areas \citep{ACTUC}. The
division of these residents into approximately equal population areas
results in dramatically different square meterage of administrative
geographic units, such as states or electorates. In a choropleth map
display, the geographic units are coloured to represent the value of the
statistic for each unit \citep{EI}. This can cause some values to be
emphasised over others, and allows choropleth map displays to
misrepresent the spatial distributions of human related statistics due
to area-size bias \citep{BCM}.

The Australian Cancer Atlas is an online, interactive display of
Australian cancer statistics in each of the Statistical Areas at Level 2
(SA2s), used by the \citet{abs2011}. The estimated standardised
incidence ratios (SIRs) of thyroid cancer for females was downloaded
from the publicly accessible Australian Cancer Atlas website and
presented in Figure \ref{fig:choro}, a choropleth map that uses colour
to display the distribution. The Australian choropleth map display draws
attention to the expanse of dark and light blue areas across the rural
communities in all states. The SA2s on the east coast around Brisbane
and in northern New South Wales stand out as more orange and red.
However, this display neglects the vast amount of Australia residents
living in the densely populated capital cities.

@@

The solutions to this visualisation problem begin with the geography.
Alternative maps can be created to shift the focus from land area and
shape to the value of the statistics \citep{ACCAC}. Each style of
cartogram applies a different transformation to the geographic areas, to
highlight the values of the statistic of interest. Alternative maps can
result in a distortion of the map space to represent features of the
distribution across the areas \citep{ACCAC} as the statistic of interest
is used to determine the cartogram layout.

Alternative mapping methods, like cartograms implemented in
\CRANpkg{cartogram}, allow increased understanding of the spatial
distribution of a variable across the population, by fairly representing
the population in each administrative area \citep{TAAM}. This
acknowledges that the number of residents can be different and also
recognises that each area, or person within it is equally important.

@@

The familiar choropleth map display is often used to allow the user to
orient themselves. However, the emphasis on the land mass diminishes the
features of the distribution in densely populated communities due to the
small size on the display \citep{ACTUC}. The unique shapes of boundaries
can be helpful for orienting users but may not contribute to their
understanding of the spatial disease distribution as many of the
communities are not visible in a choropleth display \citep{TVSSS}.

@@

There are several established alternative visualisation methods.
Rectangular cartograms \citep{ORC} and Dorling cartograms \citep{ACTUC}
implemented in \pkg{cartogram} and tile maps implemented in
\CRANpkg{tilemaps}, all use one simple shape to represent each
geographic unit. They all minimise the emphasis on the size or shape of
the geographic areas. These alternative map displays focus on the
relationship between neighbours, attempting to preserve connections, and
disregard the unique shapes of the administrative boundaries. Figure
\ref{fig:tasdisplays} shows a collection of alternative map displays,
this includes a) a contiguous cartogram, b) a non-contiguous cartogram,
and c) a Dorling cartogram.

@@

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The connected communities can be
recognised by allowing the boundaries to maintain connection in the
transformed display. The contiguous cartogram displayed in Figure
\ref{fig:tasdisplays}a draws attention to smaller geographic units when
they are rescaled according to the population \citep{DMAHP}. These new
shapes can now be coloured to represent a second variable. The algorithm
uses the geographic shape of the areas and iterates toward sizing the
areas to represent the population. This display can create twisted and
unfamiliar shapes from the geographic units as the algorithms must
satisfy the topology conditions, especially when there are communities
located geographically far from their neighbours \citep{TVSSS}.

The non-contiguous cartogram in Figure \ref{fig:tasdisplays}b also uses
the population to rescale the geographic units. Unlike the contiguous
cartogram, the SA2 areas maintain their geographic shape, but they may
not retain the connection with their neighbours. The population of the
SA2 areas is used to scale the geographic units in the non-contiguous
cartogram \citep{NAC}. The amount of background space can be meaningful
in non-contiguous cartograms \citep{ECGC}. The size disparity between
rural areas and urban areas result in reasonable display within the
south eastern city of Hobart, but these units are not visible in the
context of the Tasmania state map. Depending on the difference between
the population and geographic land size, the amount of white space can
also prevent meaningful understanding of the distribution \citep{TVSSS}.

The Dorling cartogram presents each geographic unit as a circle, the
size of the circle is scaled according to the population value of each
area \citep{ACTUC}. Figure \ref{fig:tasdisplays}c shows the Tasmania SA2
areas as an individual circle located as close as possible to the
geographic centroid location. This map draws attention to the collection
of coastal cities in Tasmania that were not apparent in Figure
\ref{fig:tasdisplays} a or b. This display also highlights that the
approximation of equal population is not easy to fulfil, as there is
some disparity in the sizes of the circles.

It is also common to treat all geographic units as equal, so that
regardless of the population or other information, to generate equal
size spatial polygons. The Dorling algorithm can be induced to do this,
to make equallly sized circles. \citet{MDAC} describes a group of
alternative maps called ``mosaic cartograms'' for tile displays. The
geography of the USA has been transformed into a square-binned tile map
manually and made available in the R package \CRANpkg{statebins}
\citep{statebins}.The algorithm described below is an automatic layout
of hexagon tiles, where a continguity constraint is relaxed. This gives
equal weight to each geographic area even if the population differs.

@@

Compared to the choropleth map display in Figure \ref{fig:choro}, the
low rates in the rural Australian communities are no longer dominating
the display. While the representative hexagons are still visible against
the black background, the much lower than average rates in rural
Australia are less noticeable.

@@

There are several key steps in the creation of the hexagon tile map as
described in the flow chart in Figure \ref{fig:sugarbagflow}. First,
derive the set of centroids from the polygons provided, then create the
grid of hexagon locations. These two processes are defined in the blue
left column of the flow chart in figure \ref{fig:sugarbagflow}. Each
centroid can then be allocated to an available hexagon location. The
steps for the allocation process are detailed in the right column of
Figure \ref{fig:sugarbagflow}. There are several filter steps to speed
the process of selecting an appropriate hexagon to represent each
geographic unit. To make tessellated plots with the hexagon allocations,
the point locations are converted into hexagon shapes.

@@

Hexagon tile maps can be useful to understand a distribution across a
collection of geographic areas. However, these maps are not easy to
create manually, especially as the number of areas increases. This
algorithm was created to automate this process, and reduce the manual
load involved in creating and implementing alternative displays. This
allows map makers and data communicators to spend their time choosing
the most effective display.

The \pkg{sugarbag} package contains a set of functions that help R users
to create a hexagon tile map. The algorithm presented in the
\pkg{sugarbag} package operates on a set of simple feature geometry
objects , known as \CRANpkg{sf} objects\citep{sf}. This package allows R
users to create \pkg{sf} objects by importing polygons stored in various
formats. Users should provide a set of polygons that define geographic
units by their administrative boundaries. The functions arrange the
geographic units in order of proximity to a set of locations provided,
such as the centre of major cities. The centroid location of each
geographic unit is used to measure the proximity. It emphasises the
major cities as population hubs, rather than emphasizing the size of
large, rural geographic units.

A user can tweak the parameters of the hexagon map using the additional
arguments to the \code{create\_hexmap} function, but these options may
affect the speed of the algorithm. The hexagon size may need adjustments
depending on the density of the population; users can provide an
appropriate hexagon size and re-run the algorithm. The buffer distance
may need to be increased if the coastal cities need to extend beyond the
geographic land mass.

@@

If a user would like to perform each of the steps of the algorithm
themselves, the necessary inputs will change for each function. For
example, the \code{allocate} function can be used directly when a user
wishes to use a set of centroids, rather than polygons.

The set of SA2 polygons for Tasmania, using the 2011 definition of SA2s,
is accessed from the \pkg{absmapsdata} package. A single column of the
data set is used to identify the unique areas. In this case, the unique
SA2 names for each SA2 have been used.

@@

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate hexagon locations
for allocation, the \code{create\_grid} function will call the
\code{create\_buffer} function. It finds the grid points needed to best
capture the set of centroids on a hexagon tile map. Reducing the buffer
can help to decrease the amount of time taken, as there will be less
points over the water to consider for each centroid allocation.

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate hexagon locations
for allocation, the \texttt{create\_grid} function will call the
\texttt{create\_buffer} function. It finds the grid points needed to
best capture the set of centroids on a hexagon tile map. Reducing the
buffer can help to decrease the amount of time taken, as there will be
less points over the water to consider for each centroid allocation.

The closest latitude row and longitude column are found for each
centroid location. Then rows and columns of centroids are divided into
20 groups. The amount of rows in each latitude group and the amount of
columns in each longitude group are used as the width of rolling
windows. The first rolling window step finds the minimum and maximum
centroid values within each of the sliding window groups of longitude
columns, and the groups of latitude rows. The second rolling window step
finds the average of the rolling minimum and maximum centroid values,
for the longitude columns and latitude rows.

@@

The set of possible hexagon grid points reduces by one after each
centroid is allocated, then only the grid points that have not yet been
selected are considered. Keeping only the available hexagon points
prevents multiple geographic units from being allocated to the same
hexagon. This is demonstrated in Figure \ref{fig:filterprocess} 2a and
2b by the black hexagons that represent the seven closest polygons to
the capital city of Hobart. As the allocation process begins for the
eighth closest centroid there are seven unavailable hexagon locations.

@@

A selection of possible hexagon locations around the centroid allows
only the closest points that are not yet assigned to be considered. The
number of possible hexagon locations to consider for a centroid is
determined by the hexagon filter. This is the maximum amount of hexagons
between the centroid and the furthest considered hexagon. It is used to
subset possible grid points to only those surrounding the polygon
centroid within an appropriate range. A smaller distance will increase
speed, but can decrease accuracy when width of the angle increases.

The algorithm creates a circle of points, by only keeping points within
a certain radial distance around the original centroid location.

The \code{width} parameter is used to take a slice of the remaining
points. The width is the amount of degrees used on either side of the
angle from the focal point to centroid location. This uses the angle
from the closest capital city, to the current centroid as seen in Figure
\ref{fig:filterprocess} 2a. This allows the spatial relationship to be
preserved, even when it is allocated to a hexagon that is further from
the focal point then the original centroid location.

@@

This work provides an algorithm called the hexagon tile map. It also
describes how this algorithm addresses the potential issues found with
contiguous, non-contiguous and Dorling cartograms when applied to
Australia. It achieves a display that effectively presents the density
of the population in small geographic areas and de-emphasises the large
rural geographic space between the densely populated capital cities. The
hexagon tile map display acknowledges that the amount of residents can
differ but each administrative area is equally important. This solution
uses equally sized areas, and maintain neighbourhood boundary
connections.

The hexagon tile map display acknowledges that the amount of residents
can differ but each administrative area is equally important. This
solution uses equally sized areas, and maintain neighbourhood boundary
connections. The \pkg{sugarbag} package for R automates the creation of
tessellated hexagon tile maps by providing an algorithm to design these
displays. The Australian application preserves spatial relationships,
and emphasises capital cities. However, similar to the choropleth map
display the tessellation does not allow the size of the hexagons to
represent another variable. The algorithm is heavily dependent on the
focal points used, as this determines the order of allocation. With
careful consideration of the choropleth map, the small geographic inner
city areas may not have been noticed by viewers, but the hexagon tile
map display emphasises them. The communities in northern Tasmania and
the Northern territory do not draw attention because of their size as in
the choropleth, but their colour is still noticeably below average when
contrasted with the hexagons in New South Wales.

Australia is used as an example, because it represents an extreme sparse
geography and small dense population areas which produces unsatisfactory
results with all existing cartogram-type algorithms. However, the
algorithm is sufficiently general to apply to other geographic regions,
and indeed has been applied to make hexagon maps of the UK, USA and
France.

There are many possible extensions to the algorithm that might be
valuable future work. The buffer allows extension beyond the furthest
centroids, but there is no mechanism to protect borders to force
centroids are placed within the geographic borders of a region
(e.g.~country or state). The algorithm is also currently limited to
representing each geographic area with only one hexagon. In the filter
step for final selection of a hexagon grid point for each centroid, a
direct angle constraint is currently implemented, but it may narrowly
miss potentially better hexagon allocation. Some other constraints, such
as a logarithmic function may help to choose refine the final allocation
to hexagons closer to the original centroid location. In a country like
Australia, the vast centre produces a hexagon map with small isolated
hexagons, even when the city centres are hugely expanded. It could be
useful to allow variable sized hexagons, primarily to increase the size
of the isolated hexagons to improve their visibility.

>>>>

\abstract{%
Spatial distributions have been presented on alternative representations
of geography, such as cartograms, for many years. In modern times,
interactivity and animation have allowed alternative displays to play a
larger role. Alternative representations have been popularised by online
news sites, and digital atlases with a focus on public consumption.
Applications are increasingly widespread, especially in the areas of
disease mapping, and election results. The algorithm presented here
creates a display that uses tessellated hexagons to represent a set of
spatial polygons, and is implemented in the R package called sugarbag.
It allocates these hexagons in a manner that preserves the spatial
relationship of the geographic units, in light of their positions to
points of interest. The display showcases spatial distributions, by
emphasising the small geographical regions that are often difficult to
locate on geographic maps.
}

@@

Over time, the population density in major cities has increased as
residents have gathered to live near urban areas \citep{ACTUC}.
Populations tend to be distributed unevenly across geographic regions.
When comparing sets of administrative geographic units such as states or
electorates, area size is seldom proportional to the population size. In
a choropleth map display, the geographic units are coloured to represent
the value of the statistic for each unit \citep{EI}. This can cause some
values to be emphasised over others, and allows choropleth map displays
to misrepresent the spatial distributions of human related statistics
due to area-size bias \citep{BCM}.

The Australian Cancer Atlas is an online, interactive display of
Australian cancer statistics in each of the Statistical Areas at Level 2
(SA2s), used by the \citet{abs2011}. The dataset of estimated
standardised incidence ratios (SIRs) of thyroid cancer for females was
downloaded from the publicly accessible Australian Cancer Atlas website
and presented in Figure \ref{fig:choro}, a choropleth map that uses
colour to display the distribution. The Australian choropleth map
display draws attention to the expanse of dark and light blue areas
across the rural communities in all states. The SA2s on the east coast
around Brisbane and in northern New South Wales stand out as more orange
and red. However, this display neglects the vast amount of Australian
residents living in the densely populated capital cities.

@@

The solutions to this visualisation problem begin with the geography.
Alternative maps can be created to shift the focus from land area and
shape to the value of the statistics \citep{ACCAC} in a cartogram
display. Cartogram methods apply different transformations to the
geographic areas, to highlight the values of the statistic of interest.
Alternative maps can result in a distortion of the map space to
represent features of the distribution across the areas \citep{ACCAC} as
the statistic of interest is used to determine the cartogram layout.

Alternative mapping methods, like cartograms implemented in
\CRANpkg{cartogram} \citep{cartogram} for R \citep{R}, promote better
understanding of the spatial distribution of a variable across the
population, by representing the population in each administrative area
fairly \citep{TAAM}. This acknowledges that the number of residents can
be different and also recognises that each area, or person within it is
equally important.

@@

Typically, chloropleth maps are used to orient the users within the
geographic context. However, the strong salience of the land mass can
hide or downplay the features of the distribution in densely populated
communities due to the small size on the display \citep{ACTUC}. The
unique shapes of boundaries can be helpful for orienting users but may
not contribute to their understanding of the spatial disease
distribution as many of the communities are not visible in a choropleth
display \citep{TVSSS}.

@@

There are several established alternative visualisation methods.
Rectangular cartograms \citep{ORC} and Dorling cartograms \citep{ACTUC}
implemented in \pkg{cartogram} and tile maps implemented in
\CRANpkg{tilemaps} \citep{tilemaps}, all use one simple shape to
represent each geographic unit. They all minimise the salience of the
size or shape of geographic areas. These alternative map displays
highlight the relationship between neighbours, preserve connections, and
discard the unique shapes of the administrative boundaries. Figure
\ref{fig:tasdisplays} shows a collection of alternative map displays,
this includes a) a contiguous cartogram, b) a non-contiguous cartogram,
and c) a Dorling cartogram.

@@

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The connectedness of the units can
be preserved by using transformations that maintain the connection
between boundaries. The contiguous cartogram displayed in Figure
\ref{fig:tasdisplays}a draws attention to smaller geographic units when
they are rescaled according to the population \citep{DMAHP}. These new
shapes can now be coloured to represent a second variable. The algorithm
uses the geographic shape of the areas and iterates toward sizing the
areas to represent the population. This display can create twisted and
unfamiliar shapes from the geographic units as the algorithms must
satisfy the topology conditions, especially when there are communities
located geographically far from their neighbours \citep{TVSSS}.

The non-contiguous cartogram in Figure \ref{fig:tasdisplays}b also uses
the population to rescale the geographic units. Unlike the contiguous
cartogram, the SA2 areas maintain their geographic shape, but they may
not retain the connection with their neighbours. The population of the
SA2 areas is used to scale the geographic units in the non-contiguous
cartogram \citep{NAC}. The amount of background space can be meaningful
in non-contiguous cartograms \citep{ECGC}. However, the relative size of
units in comparison to the reference unit can lead to scaling issues.
The size disparity between rural areas and urban areas result in
reasonable display within the south eastern city of Hobart, but these
units are not visible in the context of the Tasmania state map.
Depending on the difference between the population and geographic land
size, the amount of white space can also prevent meaningful
understanding of the distribution \citep{TVSSS}.

The Dorling cartogram presents each geographic unit as a circle, the
size of the circle is scaled according to the population value of each
area \citep{ACTUC}. Figure \ref{fig:tasdisplays}c shows the Tasmania SA2
areas as an individual circle located as close as possible to the
geographic centroid location. This map draws attention to the collection
of coastal cities in Tasmania that were not apparent in Figure
\ref{fig:tasdisplays} a or b. This display also highlights the
difference in the population of each unit, as there is some disparity in
the sizes of the circles.

Another common practice is to use equal-sized polygons and treat all
geographic units as equal. The Dorling algorithm can be induced to do
this, to make equallly sized circles. \citet{MDAC} describes a group of
alternative maps called ``mosaic cartograms'' for tile displays.
Similarly, the geography of USA has been transformed into a
square-binned tile map manually and made available in the R package
\CRANpkg{statebins} \citep{statebins}.The algorithm described below
generates an automatic layout of hexagon tiles, where a continguity
constraint is relaxed. This gives equal weight to each geographic area
even if the population differs.

@@

Compared to the choropleth map display in Figure \ref{fig:choro}, the
low rates in the rural Australian communities no longer dominate the
display. While the corresponding hexagons are still visible against the
black background, the lower rates in rural Australia are less prominent.

@@

There are several key steps in the creation of the hexagon tile map as
described in the flow chart in Figure \ref{fig:sugarbagflow}. First,
derive the set of centroids from the polygons provided, then create the
grid of hexagon locations. These two processes are described in the blue
left column of the flow chart in figure \ref{fig:sugarbagflow}. Each
centroid can then be allocated to an available hexagon location. The
steps for the allocation process are detailed in the right column of
Figure \ref{fig:sugarbagflow}. There are several filter steps to speed
up the process of selecting an appropriate hexagon to represent each
geographic unit. To make tessellated plots with the hexagon allocations,
the point locations are converted into hexagon shapes.

@@

Hexagon tile maps can be useful for visualising distributions across a
collection of geographic areas. However, these maps are not easy to
create manually, especially as the number of areas increases. This
algorithm was developed to automate the process, and reduce the workload
involved in creating and implementing alternative displays. This allows
map makers and data communicators to spend their time choosing the most
effective display.

The \pkg{sugarbag} package contains a set of functions that help R users
to create a hexagon tile map. The algorithm presented in the
\pkg{sugarbag} package operates on a set of simple feature geometry
objects , known as \CRANpkg{sf} objects \citep{sf}. This package allows
R users to create \pkg{sf} objects by importing polygons stored in
various formats. Users should provide a set of polygons that define
geographic units by their administrative boundaries. The functions
arrange the geographic units in order of proximity to a set of locations
provided, such as the centre of major cities. The centroid location of
each geographic unit is used to measure the proximity. It emphasises the
major cities as population hubs, rather than emphasizing the size of
large, rural geographic units.

The user can tweak the parameters of the hexagon map using additional
arguments to the \code{create\_hexmap} function, but these options may
affect the speed of the algorithm. The hexagon size may need adjustments
depending on the density of the population; users can provide an
appropriate hexagon size and re-run the algorithm. The buffer distance
may need to be increased if the coastal cities need to extend beyond the
geographic land mass.

@@

If a user would like to perform steps of the algorithm themselves,
additional user input will be needed for the functions that perform each
step. For example, if the user wishes to use a set of centroids, rather
than polygons, the \code{allocate} function can be used directly.

The set of SA2 polygons for Tasmania, using the 2011 definition of SA2s,
was accessed from the \pkg{absmapsdata} package \citep{absmapsdata}. A
single column of the data set is used to identify the unique areas. In
this case, the unique SA2 names for each SA2 have been used.

@@

Not all of the grid points will be used, especially if islands result in
a large grid space. To filter the grid for appropriate hexagon locations
for allocation, the \code{create\_grid} function will call the
\code{create\_buffer} function. It finds the grid points needed to best
capture the set of centroids on a hexagon tile map. Reducing the buffer
size will decrease the amount of time that the algorithm needs to run
for, as there will be less points over the water to consider for each
centroid allocation.

For each centroid location, the closest latitude row and longitude
column are found. Then rows and columns of centroids are divided into 20
groups. The number of rows in each latitude group and the number of
columns in each longitude group are used as the width of rolling
windows. The first rolling window step finds the minimum and maximum
centroid values within each of the sliding window groups of longitude
columns, and the groups of latitude rows. The second rolling window step
finds the average of the rolling minimum and maximum centroid values,
for the longitude columns and latitude rows.

@@

After each centroid is located, it is removed from the set of grid
points, and is no longer considered in the next step. Keeping only the
available hexagon points prevents multiple geographic units from being
allocated to the same hexagon. This is demonstrated in Figure
\ref{fig:filterprocess} 2a and 2b by the black hexagons that represent
the seven closest polygons to the capital city of Hobart. As the
allocation process begins for the eighth closest centroid there are
seven unavailable hexagon locations.

@@

The algorithm creates a circle of points, by only keeping points within
a certain radial distance around the original centroid location. Only
the hexagons which are close to the centroid and have not been assigned
are considered. The number of possible hexagon locations to consider for
a centroid is determined by the hexagon filter. This is the maximum
number of hexagons between the centroid and the furthest considered
hexagon. It is used to subset possible grid points to only those
surrounding the polygon centroid within an appropriate range. A smaller
distance will increase speed, but can decrease accuracy when width of
the angle increases.

The \code{width} parameter is used to take a slice of the remaining
points. The slice centres on the angle from the focal point to centroid
location. This uses the angle from the closest capital city, to the
current centroid as seen in Figure \ref{fig:filterprocess} 2a. This
allows the spatial relationship to be preserved, even when it is
allocated to a hexagon that is further from the focal point then the
original centroid location.

@@

The present work describes an algorithm called the hexagon tile map. It
also shows how this algorithm addresses the potential issues found with
contiguous, non-contiguous and Dorling cartograms when applied to the
geography of Australia. It achieves a display that effectively presents
the density of the population in small geographic areas and
de-emphasises the large rural geographic space between the densely
populated capital cities. The hexagon tile map display acknowledges that
the amount of residents can differ but each administrative area is
equally important. This solution uses equally sized areas, and maintain
neighbourhood boundary connections.

The hexagon tile map display acknowledges that the amount of residents
can differ but each administrative area is equally important. This
solution uses equally sized areas, and maintain neighbourhood boundary
connections. The \pkg{sugarbag} package for R automates the creation of
tessellated hexagon tile maps by providing an algorithm to design these
displays. The Australian application preserves spatial relationships,
and emphasises capital cities. However, similar to the choropleth map
display, the tessellation does not allow the size of the hexagons to
represent another variable. The algorithm is heavily dependent on the
location and number of focal points used, as this determines the order
of allocation. With careful consideration of the choropleth map, the
small geographic inner city areas may not have been noticed by viewers,
but the hexagon tile map display emphasises them. The communities in
northern Tasmania and the Northern territory do not draw attention
because of their size as in the choropleth, but their colour is still
noticeably below average when contrasted with the hexagons in New South
Wales.

Australia serves as a good use case for the algorithm as it represents
an extremely sparse geography and small dense population areas which
cannot be satisfactorily represented by existing cartogram-like
algorithms. However, the algorithm is sufficiently general to apply to
other geographic regions, and indeed has been applied to make hexagon
maps of the UK, USA and France.

There are many possible extensions to the algorithm that might prove
valuable in the future. The buffer allows extension beyond the furthest
centroids, but there is no mechanism to protect borders by forcing
centroids to be located within the geographic borders of a set of
regions (e.g.~country or state). The algorithm is also currently limited
to representing each geographic area with one hexagon only. In the
filter step for final selection of a hexagon grid point for each
centroid, a direct angle constraint is currently implemented, but it may
narrowly miss potentially better hexagon allocation. Some other
constraints, such as a logarithmic function may help to refine the final
allocation to hexagons closer to the original centroid location. In a
country like Australia, the vast empty interior produces a hexagon map
with many small isolated hexagons, even when the city centres are hugely
expanded. It could be useful to allow variable sized hexagons, primarily
to increase the size of the isolated hexagons to improve their
visibility.
